---
title: "EVS Data Explorer — Report"
date: "`r Sys.Date()`"
output: html_document
params:
  country:       "Overall"
  outcome:       "v72"
  outcome_label: "Child suffers when mother works"
  controls:      !r NULL
  poly_age:      1
  formula_str:   "v72 ~ age"
  data:          !r NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(ggplot2)
library(dplyr)
library(broom)
library(knitr)
library(gridExtra)

# Use Shiny data if available, otherwise load from file
if (!is.null(params$data)) {
  dat <- params$data
} else {
  dat <- readRDS("data/evs_clean.rds")
}

country_label <- params$country
outcome_label <- params$outcome_label
```

## Overview

This report was generated from the **EVS Data Explorer** Shiny app.  
It summarises results for the selections below.

| Parameter | Value |
|-----------|-------|
| Country | `r country_label` |
| Outcome | `r outcome_label` |
| Additional controls | `r if (is.null(params$controls) || length(params$controls) == 0) "None" else paste(params$controls, collapse = ", ")` |
| Age polynomial degree | `r params$poly_age` |
| Model formula | `r params$formula_str` |

---

## Exploration

### Outcome: `r outcome_label`

```{r plot-outcome, fig.height=3}
ggplot(dat, aes(x = .data[[params$outcome]])) +
  geom_histogram(bins = 4, fill = "#2c7bb6", colour = "white") +
  scale_x_continuous(breaks = 1:4) +
  labs(title = outcome_label, x = "Response (1–4)", y = "Count") +
  theme_minimal(base_size = 13)
```

### Control variables

```{r plot-controls, fig.height=3, fig.width=10}
p_age <- ggplot(dat, aes(x = age)) +
  geom_histogram(bins = 30, fill = "#1a9641", colour = "white") +
  labs(title = "Age", x = "Age", y = "Count") +
  theme_minimal(base_size = 12)

p_edu <- ggplot(dat, aes(x = edu)) +
  geom_bar(fill = "#d7191c") +
  labs(title = "Education", x = "Level", y = "Count") +
  theme_minimal(base_size = 12)

p_sex <- ggplot(dat, aes(x = sex)) +
  geom_bar(fill = "#fdae61") +
  labs(title = "Sex", x = "Sex", y = "Count") +
  theme_minimal(base_size = 12)

gridExtra::grid.arrange(p_age, p_edu, p_sex, ncol = 3)
```

---

## Regression

**Formula:** `r params$formula_str`

```{r regression}
fit <- lm(as.formula(params$formula_str), data = dat)
```

### Coefficients

```{r coef-table}
tidy(fit, conf.int = TRUE) |>
  mutate(across(where(is.numeric), \(x) round(x, 4))) |>
  rename(
    Term = term,
    Estimate = estimate,
    `Std. Error` = std.error,
    Statistic = statistic,
    `p-value` = p.value,
    `CI low` = conf.low,
    `CI high` = conf.high
  ) |>
  kable()
```

### Predicted values vs. Residuals

```{r resid-plot, fig.height=4}
df_resid <- data.frame(
  fitted    = fitted(fit),
  residuals = residuals(fit)
)

ggplot(df_resid, aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.3, colour = "#2c7bb6") +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "red") +
  geom_smooth(method = "loess", se = FALSE, colour = "#d7191c") +
  labs(
    title = "Predicted vs. Residuals",
    x = "Fitted values",
    y = "Residuals"
  ) +
  theme_minimal(base_size = 13)
```
